Concurrent Programming in Harmony

### 2020-09-27 book/paper.tex
Fixed Peterson's proof

To see that $\mathcal{I}_p(0) \land \mathcal{I}_p(1)$ is, in fact, an inductive invariant, first note that
it holds in the initial state, because in the initial state no process
is in the critical section.
Without loss of generality, suppose $i=0$ (a benefit from the fact that the algorithm is
symmetric for both processes).  We still have to show that if we are in a state
in which $\mathcal{I}_p(0)$ holds, then any step will result in a state in which
$\mathcal{I}_p(0)$ still holds.
If $\mathcal{I}_p(0)$ holds, process~0 is at label \texttt{@cs}.  If process~0
were to take a step, then in the next state process~0 would be no longer
at that label and $\mathcal{I}_p(0)$ would hold trivially over the next state.
Therefore we only need to consider a step by process~1.

From $\mathcal{I}_p(0)$ we know that one of the following three cases must hold before
process~1 takes a step:
\begin{enumerate}
\item \textit{flags}[1] = \texttt{False};
\item \textit{turn} = 0;
\item process~1 is at label \texttt{@gate}.
\end{enumerate}

Let us consider each of these cases.
We have to show that if process~1 takes a step, then one of those
cases must hold after the step.
In the first case, if process~1 takes a step, there are two possibilities:
either $flags[1]$ will still be \texttt{False} (in which case the first case
continues to hold), or $flags[1]$ will be \texttt{True}
and process~1 will be at label \texttt{@gate} (in which case the third case
will hold).
We know that process~1 never sets \textit{turn} to 1, so
if the second case holds before the step, it will also hold after the step.
Finally, if process~1 is at label \texttt{@gate} before the step, then after
the step \textit{turn} will equal 0, and therefore the second case will hold
after the step.  qed.

>>>>

To see that $\mathcal{I}_p(0) \land \mathcal{I}_p(1)$ is, in fact, an inductive invariant, first note that
it holds in the initial state, because in the initial state no process
is in the critical section.
Without loss of generality, suppose $i=0$ (a benefit from the fact that the algorithm is
symmetric for both processes).  We still have to show that if we are in a state
in which $\mathcal{I}_p(0)$ holds, then any step will result in a state in which
$\mathcal{I}_p(0)$ still holds.

First consider the case that process~0 is at label \texttt{@cs}.  If process~0
were to take a step, then in the next state process~0 would be no longer
at that label and $\mathcal{I}_p(0)$ would hold trivially over the next state.
Therefore we only need to consider a step by process~1.
%
From $\mathcal{I}_p(0)$ we know that one of the following three cases must hold before
process~1 takes a step:
\begin{enumerate}
\item \textit{flags}[1] = \texttt{False};
\item \textit{turn} = 0;
\item process~1 is at label \texttt{@gate}.
\end{enumerate}

Let us consider each of these cases.
We have to show that if process~1 takes a step, then one of those
cases must hold after the step.
In the first case, if process~1 takes a step, there are two possibilities:
either $flags[1]$ will still be \texttt{False} (in which case the first case
continues to hold), or $flags[1]$ will be \texttt{True}
and process~1 will be at label \texttt{@gate} (in which case the third case
will hold).
We know that process~1 never sets \textit{turn} to 1, so
if the second case holds before the step, it will also hold after the step.
Finally, if process~1 is at label \texttt{@gate} before the step, then after
the step \textit{turn} will equal 0, and therefore the second case will hold
after the step.

Now consider the case where process~0 is not in the critical section,
and therefore $\mathcal{I}_p(0)$ holds trivially because false implies
anything.
There are three cases to consider:
\begin{enumerate}
\item Process~1 takes a step.  But then process~0 is still not in the critical
section and $\mathcal{I}_p(0)$ continues to hold;
\item Process~0 takes a step but still is not in the critical section.
Then again $\mathcal{I}_p(0)$ continues to hold.
\item Process~0 takes a step and ends up in the critical section.
Because process~0 entered the critical section, we know that
\textit{flags}[1] = \texttt{False} or \textit{turn} == 0 because
of the \texttt{await} condition.
And hence $\mathcal{I}_p(0)$ continues to hold in that case as well.
\end{enumerate}
